---
title: "MxIF example Processing"
output: html_notebook
---


This example assumes that imagin data was processed and segmented using the Matlab scripts.

To use the MxIF packge, the libraries below should be installed.

Load libraries and set the directory of the processed data for an individual slide:
```{r}
library(tiff)
library(ggplot2)
library(png)
library(Rtsne)
library(magick)
library(reshape)
library(jpeg)
library(plyr)
library(mclust)

library(MxIF)

#use your directory here
DIRECTORY='/Users/etmckinley/Dropbox (VUMC)/Research/Projects/GE Multiplex/R Code/MxIF/Example Data'
list.dirs(DIRECTORY)
```

Get the file names for Novlp segmentation image, the final cell segmentation image defining cell IDs, and the Statistics file for each position
```{r}
#
Novlp.spots=list.files(path = paste(DIRECTORY, 'Novlp', sep="/"))
Labels.spots=list.files(path = paste(DIRECTORY, 'CellSegFinal', sep="/"))
Stats.spots=list.files(path = paste(DIRECTORY, 'PosStats', sep="/"))
```

Example filtering and tSNE map using only the first position. A loop through all positions is included below.
```{r}
i=1
#Read in the stats data
posData=read.csv(paste(DIRECTORY, 'PosStats', Stats.spots[i], sep="/"), header=TRUE) #7480 cells using example data

#Filter out cells by size based upon the default parameters
posData=MxIF.cell.filter(posData) #4974 cells using example data

#Set the markers to run tSNE on, this example only uses Median Cell Values
markers=seq(from = 9, to = ncol(posData), by = 4)

#Find the column containing DAPI and exclude from the tSNE calcutions
dapiInd=as.numeric(grep("Median_Cell_DAPI", names(posData), value = FALSE))
markers=markers[!markers %in% dapiInd]

#run tSNE using the default parameters for the selected markers
posData=MxIF.tsne(posData, markers)

#diplay tSNE plot for Beta-Catenin
MxIF.tsne.plot(posData,9)

#save animated gif
MxIF.tsne.animate(posData, markers, paste(DIRECTORY,"animated_gif.gif", sep="/"))
```

To view the cells filtered
```{r}
library(tiff)
library(png)

#Read in the individual CellSegImage (LABELS) and NOVLP images 
LABELS=readTIFF(paste(DIRECTORY, 'CellSegFinal' , Labels.spots[i] ,sep="/"), as.is=TRUE)
NOVLP=readPNG(paste(DIRECTORY, 'Novlp', Novlp.spots[i] , sep="/"))
NOVLP=MxIF.cell.filter.view(posData, LABELS, NOVLP)
writePNG(NOVLP, target=paste(DIRECTORY, 'Cells_filtered.png',sep="/"))
```

To overlay selected cells on an image
```{r}
# Read in labels image and AFremoved image
LABELS=readTIFF(paste(DIRECTORY, 'CellSegFinal' , Labels.spots[i] ,sep="/"), as.is=TRUE)
AFRemoved=readJPEG(paste(DIRECTORY, 'BCAT_AFRemoved_spot_000.jpg', sep="/") )


#select the top 10 highest B-catenin expressing cells and overlay
column_select=9
top10=posData[order(-posData[,column_select]),]
top10=top10[1:10,]
top10=as.array(top10$ID)
 
Out=MxIF.cell.ID.view(AFRemoved, LABELS, top10)
 
writeJPEG(Out, target=paste(DIRECTORY, "B-cat_top10.jpg",sep="/"))

```

Using model-based binary deconvolution thresholding and creating "Death Star" Plots
```{r}
#Binarization on all cells while retaining Cell ID and plotting histograms
BinaryData=MxIF.deconvolution(posData, c(9,13,17,21,29), 'SlideName', DIRECTORY, plots=TRUE)

#Death star plot
MxIF.death.star.plot(BinaryData, 3:7, DIRECTORY)

#Binarization of highest 20th perentile of a marker
BinaryData20=MxIF.deconvolution(posData, c(9,13,17,21,29), 'SlideName2', DIRECTORY, selectedMarker = names(posData)[9], quant=.8, plots=TRUE)

#Death Star Plot
MxIF.death.star.plot(BinaryData20, 3:7, DIRECTORY)


```


For processing multiple positions at a time.
```{r}
library(tiff)
library(ggplot2)
library(png)
library(Rtsne)
library(magick)
library(reshape)
library(MxIF)
library(jpeg)

DIRECTORY='/Users/etmckinley/Dropbox (VUMC)/Research/Projects/GE Multiplex/R Code/MxIF/Example Data'

#get filenames of spots
Novlp.spots=list.files(path = paste(DIRECTORY, 'Novlp', sep="/"))
Labels.spots=list.files(path = paste(DIRECTORY, 'CellSegFinal', sep="/"))
Stats.spots=list.files(path = paste(DIRECTORY, 'PosStats', sep="/"))

#create folders for storing tSNE maps, and filtered data if they don't already exist
dir.create(paste(DIRECTORY, 'tSNE', sep="/"), showWarnings = FALSE)
dir.create(paste(DIRECTORY, 'PosStats_filt', sep="/"), showWarnings = FALSE)
dir.create(paste(DIRECTORY, 'Novlp_filt', sep="/"), showWarnings = FALSE)

#loop through each position
for (i in 1:length(Stats.spots)){
  
    #check if file exists for each position
   if (file.exists(paste(DIRECTORY, 'PosStats_filt', Stats.spots[i], sep="/"))){
      posData=read.csv(paste(DIRECTORY, 'PosStats_filt', Stats.spots[i], sep="/"), header=TRUE)
  }else {

       #read data for each position
      posData=read.csv(paste(DIRECTORY, 'PosStats', Stats.spots[i], sep="/"), header=TRUE) 
    
      #Filter out cells by size based upon the default parameters
      posData=MxIF.cell.filter(posData) #4974 cells using example data
  
      #Set the markers to run tSNE on, this example only uses Median Cell Values
      markers=seq(from = 9, to = ncol(posData), by = 4)

      #Find the column containing DAPI and exclude from the tSNE calcutions
      dapiInd=as.numeric(grep("Median_Cell_DAPI", names(posData), value = FALSE))
      markers=markers[!markers %in% dapiInd]
      
      #run tSNE using the default parameters for the selected markers
      posData=MxIF.tsne(posData, markers)
    
      #write filtered cells 
      write.csv(posData, paste(DIRECTORY, 'PosStats_filt', Stats.spots[i], sep="/"), row.names = FALSE)
  }
  
  ###
  ###write out tSNE maps for each position and an animated gif, comment out if not desired
  ###
  
  #get the string for the position 
  pos=gsub("Novlp_", "", Novlp.spots[i])
  pos=gsub(".png", "", pos)
  
  #select markers to output
  markers=c(seq(from = 9, to = ncol(posData)-2, by = 4),ncol(posData)-1,ncol(posData))
  
  #loop through each marker, create and save tSNE maps
  for (j in markers){
    p=MxIF.tsne.plot(posData,j)
    ggsave(filename=paste(DIRECTORY, '/tSNE/',pos, "_", names(posData)[j], ".png", sep=""), plot=p, height=6, width=6, bg = "transparent")
}

  # create and save animated gif
  MxIF.tsne.animate(posData, markers, paste(DIRECTORY, '/tSNE/',pos, "_animation.gif", sep=""))

  ###
  ##generate visualization of filtered cells from NOVLP segmentation image, comment out if not desired
  ###
  
  #load the labels image and the NOVLP segmenation image
  LABELS=readTIFF(paste(DIRECTORY, 'CellSegFinal' , Labels.spots[i] ,sep="/"), as.is=TRUE)
  NOVLP=readPNG(paste(DIRECTORY, 'Novlp', Novlp.spots[i] , sep="/"))

  #Run function to create cell filtered visualization
  NOVLP=MxIF.cell.filter.view(posData, LABELS, NOVLP)
  
  #modify filename and save image
  filtFile=gsub("Novlp", "Novlp_filt", Novlp.spots[i])
  writePNG(NOVLP, target=paste(DIRECTORY, 'Novlp_filt', filtFile,sep="/")) 

}
```

